A tutorial of Bayesian beta regressions with brms in R

### 2024-12-06 manuscript/manuscript.qmd
edits based on Paul

We will use brms to fit Bayesian beta regressions. [XXX why Bayesian]. 
The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.^[Footnote about Gelman's terminology for random effects.] The categorical predictors vowel and C2 are coded using indexing rather than traditional R contrasts (in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts does not impact the model posterior predictions, but it makes it easier to specify priors when users do so). For pedagogical simplicity, the model will use the default priors, but note that in real data analyses contexts, priors should be specified by the user. I refer the readers to XXX.

@@

The `summary()` function prints the full model summary. For conciseness, we will use the `fixef()` function which prints the regression coefficients of the constant (aka fixed) terms. The full summary with an explanation of each part can be found in XXX. For each coefficient in the model, `fixef()` prints the name of the coefficient, the mean estimate, the estimate error and the lower and upper limits of a Bayesian Credible interval (CrI). Here, we print an 80% CrI.^[There is nothing special about 95% within Bayesian inference and instead experts recommend to check and report a variety of CrIs. For simplicity, we will use 80% CrIs in this case study.] In Bayesian inference, an 80% CrI indicates the range of values within which the true estimate falls at 80% probability or confidence. We round all values to the nearest 2 digits for clarity.

@@

The coefficients of a beta regression are estimated on the log-odds scale, as in Bernoulli/binomial (aka logistic) regressions. From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds [$\beta$ = 0.8, SD = 0.06]. Log-odds can be converted to odd ratios by exponentiating the value: 0.01-0.15 log odds correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an increase of voicing of 1 to 16% for every increase of one syllable per second. Since this is an 80% CrI, we can be 80% confident that the true effect of speech rate is between 1-16% increase of voicing proportion.

>>>>

We will use brms to fit Bayesian beta regressions. [XXX why Bayesian]. 
The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.^[Footnote about Gelman's terminology for random effects.] The categorical predictors vowel and C2 are coded using indexing rather than traditional R contrasts: in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts makes it easier to specify priors. For pedagogical simplicity, the model will use the default priors, but note that in real data analyses contexts, priors should be specified by the user. I refer the readers to XXX.

@@

The `summary()` function prints the full model summary. For conciseness, we will use the `fixef()` function which prints the regression coefficients. The full summary with an explanation of each part can be found in XXX. For each coefficient in the model, `fixef()` prints the name of the coefficient, the mean estimate, the estimate error and the lower and upper limits of a Bayesian Credible interval (CrI). Here, we print an 80% CrI. There is nothing special about 95% CrI within Bayesian inference and instead experts recommend to check and report a variety of CrIs. Bayesian CrIs indicate that at a certain probability levels the "true" estimate lies within that interval: so, for example, a 90% CrI [A, B] indicates that there is a 90% probability that the "true" estimate is between A and B. Different probability levels correspond to different levels of confidence: the higher the probability the higher the confidence (always conditional on data and model). Obtaining CrIs at different probability levels allows researchers to make more fine-grained inferential statements than the frequentist significance dichotomy affords. For simplicity of exposition, we will use 80% CrIs in this case study but we strongly recommend researchers to always obtain CrIs at different levels of probability and base their inferences on all and not one in particular. To reiterate, in Bayesian inference, an 80% CrI indicates the range of values within which the true estimate falls at 80% probability or confidence. We round all values to the nearest 2 digits for clarity.

@@

The coefficients of a beta regression are estimated on the log-odds scale, as in Bernoulli/binomial (aka logistic) regressions. From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds [$\beta$ = 0.8, SD = 0.06]. Log-odds can be converted to odd ratios by exponentiating the value: 0.01-0.15 log odds correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an increase of voicing of 1 to 16% for every increase of one syllable per second. Since this is an 80% CrI, we can be 80% confident that the true effect of speech rate is between 1-16% increase of voicing proportion, conditional on the data and model. Note that transforming measures this way is appropriate *only* with quantile-based measures (like CrIs) but not with moments like the mean and standard deviation: to correctly get mean and SDs in the transformed scale, you must first extract the posterior draws (see below), convert them and then take moments such as mean and SD (for a more detailed explanation, see XXX).

### 2024-12-06 manuscript/manuscript.tex
edits based on Paul

We will use brms to fit Bayesian beta regressions. {[}XXX why
Bayesian{]}. The model has voicing proportion as the outcome variable
and the following terms: an interaction between vowel (/i, e, a, ɔ, u/)
and second consonant C2 (/k, t, p/), centred speech rate (number of
syllables per second); as varying (aka random) terms, by-speaker varying
coefficients for the vowel/consonant interaction and for centred speech
rate.\footnote{Footnote about Gelman's terminology for random effects.}
The categorical predictors vowel and C2 are coded using indexing rather
than traditional R contrasts (in R, this corresponds to suppressing the
model's intercept with the \texttt{0\ +} syntax; using indexing instead
of contrasts does not impact the model posterior predictions, but it
makes it easier to specify priors when users do so). For pedagogical
simplicity, the model will use the default priors, but note that in real
data analyses contexts, priors should be specified by the user. I refer
the readers to XXX.

@@

The \texttt{summary()} function prints the full model summary. For
conciseness, we will use the \texttt{fixef()} function which prints the
regression coefficients of the constant (aka fixed) terms. The full
summary with an explanation of each part can be found in XXX. For each
coefficient in the model, \texttt{fixef()} prints the name of the
coefficient, the mean estimate, the estimate error and the lower and
upper limits of a Bayesian Credible interval (CrI). Here, we print an
80\% CrI.\footnote{There is nothing special about 95\% within Bayesian
  inference and instead experts recommend to check and report a variety
  of CrIs. For simplicity, we will use 80\% CrIs in this case study.} In
Bayesian inference, an 80\% CrI indicates the range of values within
which the true estimate falls at 80\% probability or confidence. We
round all values to the nearest 2 digits for clarity.

@@

The coefficients of a beta regression are estimated on the log-odds
scale, as in Bernoulli/binomial (aka logistic) regressions. From the
summary, we see that speech rate (number of syllables per second) has a
positive effect on voicing proportion: the 80\% CrI is between 0.01 and
0.15 log-odds {[}\(\beta\) = 0.8, SD = 0.06{]}. Log-odds can be
converted to odd ratios by exponentiating the value: 0.01-0.15 log odds
correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an
increase of voicing of 1 to 16\% for every increase of one syllable per
second. Since this is an 80\% CrI, we can be 80\% confident that the
true effect of speech rate is between 1-16\% increase of voicing
proportion.

>>>>

We will use brms to fit Bayesian beta regressions. {[}XXX why
Bayesian{]}. The model has voicing proportion as the outcome variable
and the following terms: an interaction between vowel (/i, e, a, ɔ, u/)
and second consonant C2 (/k, t, p/), centred speech rate (number of
syllables per second); as varying (aka random) terms, by-speaker varying
coefficients for the vowel/consonant interaction and for centred speech
rate.\footnote{Footnote about Gelman's terminology for random effects.}
The categorical predictors vowel and C2 are coded using indexing rather
than traditional R contrasts: in R, this corresponds to suppressing the
model's intercept with the \texttt{0\ +} syntax; using indexing instead
of contrasts makes it easier to specify priors. For pedagogical
simplicity, the model will use the default priors, but note that in real
data analyses contexts, priors should be specified by the user. I refer
the readers to XXX.

@@

The \texttt{summary()} function prints the full model summary. For
conciseness, we will use the \texttt{fixef()} function which prints the
regression coefficients. The full summary with an explanation of each
part can be found in XXX. For each coefficient in the model,
\texttt{fixef()} prints the name of the coefficient, the mean estimate,
the estimate error and the lower and upper limits of a Bayesian Credible
interval (CrI). Here, we print an 80\% CrI. There is nothing special
about 95\% CrI within Bayesian inference and instead experts recommend
to check and report a variety of CrIs. Bayesian CrIs indicate that at a
certain probability levels the ``true'' estimate lies within that
interval: so, for example, a 90\% CrI {[}A, B{]} indicates that there is
a 90\% probability that the ``true'' estimate is between A and B.
Different probability levels correspond to different levels of
confidence: the higher the probability the higher the confidence (always
conditional on data and model). Obtaining CrIs at different probability
levels allows researchers to make more fine-grained inferential
statements than the frequentist significance dichotomy affords. For
simplicity of exposition, we will use 80\% CrIs in this case study but
we strongly recommend researchers to always obtain CrIs at different
levels of probability and base their inferences on all and not one in
particular. To reiterate, in Bayesian inference, an 80\% CrI indicates
the range of values within which the true estimate falls at 80\%
probability or confidence. We round all values to the nearest 2 digits
for clarity.

@@

The coefficients of a beta regression are estimated on the log-odds
scale, as in Bernoulli/binomial (aka logistic) regressions. From the
summary, we see that speech rate (number of syllables per second) has a
positive effect on voicing proportion: the 80\% CrI is between 0.01 and
0.15 log-odds {[}\(\beta\) = 0.8, SD = 0.06{]}. Log-odds can be
converted to odd ratios by exponentiating the value: 0.01-0.15 log odds
correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an
increase of voicing of 1 to 16\% for every increase of one syllable per
second. Since this is an 80\% CrI, we can be 80\% confident that the
true effect of speech rate is between 1-16\% increase of voicing
proportion, conditional on the data and model. Note that transforming
measures this way is appropriate \emph{only} with quantile-based
measures (like CrIs) but not with moments like the mean and standard
deviation: to correctly get mean and SDs in the transformed scale, you
must first extract the posterior draws (see below), convert them and
then take moments such as mean and SD (for a more detailed explanation,
see XXX).
